### 2021-11-03 preregistration_changes.Rmd

---
title: "Items: English"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

The following changes were introduced after the preregistration.

## Covariates

When reading the documentation of the data, I thought the following variables were varying, when in fact they were assessed only at one wave, namely the first wave:

- Employment status

## To-Do

- Analyses
	- Missingness
		- Impute missing values
		- Determine amount of missingness
		- Replace values where I know they're 0
	- Add factor scores
	- CFA
		- Add results