### 2021-11-05 index.Rmd
- Status of paper when submitted to ICA

---
title: "No meaningful effets of COVID-related social media use and well-being"
---

On this website, you can find the companion material for the paper _No meaningful effets of COVID-related social media use and well-being_.

You can find the following information here:

- The paper, which you can read [in your browser](manuscript.html) or download as [pdf](manuscript.pdf). 
- The detailed code and summaries of the [analyses](analyses.html)
- A list of the [items translated into English](items_english.html)
- A list of the [original items in German](items_german.html)
- The [preregistration](preregistration.html) 
- A list where I needed to [deviate](preregistration_changes.Rmd) from the preregistration.

This website is produced directly from the project's [github repository](https://github.com/XMtRa/Austrian_Corona_Panel_Project). You can download the material and the data, file an issue, or submit a pull request.